---
title: "Lecture 3"
author: "Andrea Boskovic"
date: '2022-06-29'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Reminders

1. Short lab 1 due at 11:59 PM tonight
2. Lab 1 will be open at 3PM today and due on 07/11 at 11:59 PM
3. No class on Monday (Happy Fourth of July!)
4. My OH are tomorrow 8-9 AM on Zoom

# Back to Lecture 1

## How do you make a table/make it look nice in R?

```{r}
# install.packages("knitr")
# install.packages("kableExtra")
library(knitr)
library(kableExtra)
```

```{r}
my_matrix <- diag(3)
my_matrix

rownames(my_matrix) <- c("First Row", "Second Row", "Third Row")
colnames(my_matrix) <- c("Column 1", "Column 2", "Column 3")
```

```{r}
kable_styling(kable(my_matrix))
my_matrix
```

The `kableExtra` package is a really useful tool for making tables and will be useful in your projects.

Look how much nicer the `kableExtra` output is than just printing `my_matrix`. 

Here's a link to explore what else this package can do: https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html

## Lists

```{r}
my_list <- list("some_numbers" = 1:5,
                "some_characters" = c("a", "b", "c"),
                "a_matrix" = diag(2))
my_list
```

**Important**: In R, the dollar sign ($) is used to refer to list items AND to column names.

### Accessing list elements

```{r}
# You can access list elements in three different ways
my_list$some_numbers
my_list[["some_numbers"]]
my_list[[1]]
```

### How do you subset a list?

Subsetting a list means accessing just part of it. Our whole list is the contents of my_list:

```{r}
my_list
```

But what if I only want to see `some_numbers` and `a_matrix`? How do I do that? This is called *subsetting*.

```{r}
my_list[c(1,3)]
my_list[c("some_numbers", "a_matrix")]
```

**Solution**: Make a vector of the list element positions that you want or the list element names.

### Adding to a list

What if I want to add another element to `my_matrix`?

```{r}
my_list$STAT302 <- "my favorite class"
my_list$good_question <- c(1, "jack", TRUE, 5000)
```

You can access the names of all your list elements with 

```{r}
names(my_list)
```

## Data Frames

```{r}
my_data <- data.frame("var1" = 1:3,
                      "var2" = c("a", "b", "c"),
                      "var3" = c(TRUE, FALSE, TRUE))
my_data
kable_styling(kable(my_data))
```

```{r}
my_matrix
my_matrix2 <- as.data.frame(my_matrix)
#typeof(my_matrix)
#typeof(my_matrix2)
```

### Subsetting Data Frames

```{r}
# Accessing a column
my_data$var1
my_data[c("var1", "var2")]

# Things can get fishy depending on how you try to extract a column
my_data["var1"]
my_data$var1
my_data[["var1"]]
```

### Adding a column

```{r}
my_data$STAT302 <- c("coding", "is", "fun")
#my_data$test <- c(1, 2, 3, 4) 
```

```{r}
# Preview into the tidyverse
my_data <- my_data %>%
  mutate(var1x2 = var1*2)
```

### Add row names

```{r}
rownames(my_data) <- c("Obs 1", "Obs 2", "Obs 3")
kable_styling(kable(my_data))
```

## Coding Style

```{r}
# my_data %>%
#   filter(var1 > 1) %>%
#   mutate(var1x3 = var1 * 3) %>%
#   group_by()
```

Addins -> Styler -> Style Selection is a really useful tool to make your code look pretty!

```{r}
AddValues <- function(x, y) {
  z <- 4
  return(x + y)
  z
}
AddValues(x = 5, y = 3)
```

Always use return!! Try taking it out in the previous chunk and see what happens!

```{r}
#first_day_of_the_month<-sdkfjdaaajjjjjjjjjjjjjjjjjjjasdoifjavdoifbjoisjdfoiajsdiovajadfjhakjsdc
```

Be careful of the margins!

```{r}
a <- 2; b <- 3 # bad practice


# good practice
a <- 2
b <- 3
```

