---
title: "Lecture 7"
author: "Andrea Boskovic"
date: '2022-07-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gapminder)
```

# Reminders

1. Lab 2 due tonight at 11:59 PM - let me know if you need an extension
2. Short lab 2 (optional/extra credit) will be assigned probably tonight
3. No class Wednesday 07/20
4. Office Hours this week:
      * 07/19 @ 1-2PM and 3-4PM
      * 07/21 @ 8-9AM
5. You should be working on Project 1 (due 1 week from today - 07/25 at 11:59 PM)

# Data Manipulation (end of Lecture 3)

Tidyverse loads a bunch of useful packages. One of them, the most important for data manipulation, is `dplyr`.

`dplyr` contains a few functions that are going to allow you to manipulate data.

The main important ones are:

1. `head()` and `tail()`: first and last observations in a dataset, respectively
2. `filter()`: allows you to select *rows* of a dataset
3. `select()`: allows you to select *columns* of a dataset
4. `mutate()`: allows you to make new columns based on existing columns
5. `group_by()` and `summarize()`: these are almost always used together. `group_by()` lets you choose a variable to group by and summarize calculates something for each category/group you chose.

Let's do some examples with the Gapminder dataset

```{r}
data(gapminder)
```

1. `head()` and `tail()`

Let's say I want to see the beginning and end of the dataset

```{r}
# gets you first 10 obs
head(gapminder, n = 10)
gapminder %>%
  head(n = 10)

# gets you last 10 obs
tail(gapminder, n = 10)
gapminder %>%
  tail(n = 10)
```

### Add in ordering

Suppose I want to order the dataset by countries with the highest life expectancy. A function that can do that is called `arrange()`. By default, it goes in ascending order of the variable you "arrange" by. If you want the other way (descending order), you can use `arrange()` and `desc()` together. 

```{r}
# better solution
gapminder %>%
  arrange(desc(lifeExp)) %>%
  head(n = 10)

# okay solution
gapminder %>%
  arrange(lifeExp) %>%
  tail(n = 10)
```

2. `filter()`

`filter()` basically lets us choose a range or a value for a variable, and takes out observations from the data that don't fit the criteria we give it. Think of it as "give me observations where ____ is true."

```{r}
# rows where country is Afghanistan
gapminder %>%
  filter(country == "Afghanistan")

# rows where year is between 2000 and 2015
gapminder %>%
  filter(year <= 2015, year >= 2000)

# you can filter on multiple different columns and separate with a comma
gapminder %>%
  filter(year <= 2015, year >= 2000,
         country == "Albania")

# you can look for a bunch of different countries
three_countries <- gapminder %>%
  filter(country %in% c("China", "Poland", "Algeria"))
three_countries
```

3. `select()` 

This lets you pair down your data by choosing columns that you want to see.

Suppose I want to show someone, in a table, Algeria's life expectancy over time.

```{r}
gapminder %>%
  filter(country == "Algeria") %>%
  select(lifeExp, year)
```

Sometimes, `select()` might not work because it's also part of another package (so you might have a conflict). In these cases, Google will tell you what to do, and the solution it'll give you is to instead use `dplyr::select()`.

### How are `select()` and `filter()` different?

Summary: 

* `select()` pairs down columns, but does not affect rows
* `filter()` pairs down rows, but does not affect columns

```{r}
gapminder %>% 
  select(country, year, pop)

gapminder %>%
  filter(gdpPercap > 750)
```

4. `mutate()`

`mutate()` lets you make new columns or change existing columns based on information from your data.

Let's say I'm interested in categorizing life expectancy into "low", "medium" and "high."

We'll say a life expectancy below 48 is low, 49-70 is medium, and above 71 is high.

```{r}
gapminder <- gapminder %>%
  mutate(exp_category = case_when(lifeExp <= 48 ~ "low",
                                  lifeExp > 48 & lifeExp <= 70 ~ "medium",
                                  lifeExp > 70 ~ "high"))
gapminder
```

Now, what if I want to say years before 2000 are old and years after (and including) 2000 are recent.

```{r}
gapminder <- gapminder %>%
  mutate(year_category = ifelse(year < 2000, "old", "recent"))
gapminder
```

5. `group_by()` and `summarize()`

These two functions are almost always used together. Generally, `group_by()` will go first, and `summarize()` will come right after it. That means that you won't see `group_by()` then `mutate()` then `summarize()` - it usually just goes `group_by() %>% summarize()`. 

Let's say I want to find the average life expectancy per country and arrange it from highest to lowest. I only want to see the top 5.

```{r}
gapminder %>%
  group_by(country) %>% # "per country"
  summarize(avg_lifeExp = mean(lifeExp)) %>%
  arrange(desc(avg_lifeExp)) %>%
  head(n = 5)
```

What if we wanted to find median GDP per continent?

```{r}
gapminder %>%
  group_by(continent) %>%
  summarize(med_GDP = median(gdpPercap)) %>%
  arrange(desc(med_GDP))
```

