---
title: "Lecture 8"
author: "Andrea Boskovic"
date: '2022-07-25'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Reminders

1. Project 1 due Wednesday at 11:59 PM
2. Optional/Extra Credit Short Lab 2 due tonight at 11:59 PM
3. There will be no short lab 3 (ignore that on the schedule - I will change it)

# Wrapping up the Tidyverse

## `pivot_longer()`

Let's use the `relig_income` dataset

```{r}
data("relig_income")
head(relig_income)
```

To put this in tidy format, we want to make the dataset longer. To do that, we can use a function in `tidyverse` called `pivot_longer()`.

In general, to make your dataset longer:

pivot_longer(name_of_data, cols = __cols_to_move_, 
                           names_to = __name of new column for "cols"_, 
                           values_to = __what to call the value column_)

```{r}
tidy_religion <- pivot_longer(relig_income,
  cols = c(
    "<$10k", "$10-20k", "$20-30k", "$30-40k", "$40-50k",
    "$50-75k", "$75-100k", "$100-150k", ">150k",
    "Don't know/refused"
  ),
  names_to = "income_group",
  values_to = "count"
)
```

## `pivot_wider()`

What if our data is long and we need to make it wide?

```{r}
data("fish_encounters")
head(fish_encounters)
```

We want to give each unique observation in the station column its own column.

In general, using `pivot_wider()` looks like this:

pivot_wider(data_name, 
            names_from = __where to pull the new column names from_,
            values_from = __where to get the values that go under each column_)

```{r}
tidy_fish <- pivot_wider(fish_encounters,
            names_from = "station",
            values_from = "seen")
```
