---
title: "lecture 4"
author: "Andrea Boskovic"
date: '2022-07-06'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Control Flow

1. `if` statements

What is an if statement?

```{r}
x <- 1

if(x > 0) {
  paste0("The value of x you put in is ", x)
}
```

```{r}
x <- -1

if(x > 0) {
  paste0("The value of x you put in is ", x)
}
```

2. `else` statements

What is an else statement?

In this case, our if statement didn't return anything. 

```{r}
x <- -1

if(x > 0) {
  paste0("The value of x you put in is ", x)
} else {
  paste0("the condition evaluated to false, but we have an else statement")
}
```

3. `else if` statements

```{r}
x <- -1

if(x > 0) {
  paste0("The value of x you put in is ", x)
} else if(x < -1) {
  paste0("hi")
} else {
  paste0("hello!")
}
```

Your turn:  Suppose we want to check if x is greater than 5 and print out the answer. What should the condition be?

```{r}
x <- 6

if(x > 5) {
  print("x is greater than 5.")
} else {
  print("x is less than or equal to 5.")
}
```

# Loops

1. `for` loops

What is it?

Allows you to iterate over a vector or a set of values until the values are exhausted.

```{r}
for(i in 1:10) {
  print(i)
}
```


```{r}
for(i in 1:10) {
  
  for(j in 1:10) {
    print(j)
  }
  
}
```

2. `while` loops

Does the same thing as a `for` loop but in a slightly messier way

```{r}
x <- 0
while(x < 5) {
  
  print(paste("x is equal to", x))
  x <- x + 1
  
}
```

## String inputs

```{r}
string_vec <- c("a", "b", "c")

for(letter in string_vec) {
  print(letter)
}
```

## Nested Loops

```{r}
for(i in 1:3) {
  
  for(j in 1:2) {
    
    print(i * j)
    
  }
  
}
```

## Filling in Vectors using Loops

```{r}
# the vector i want to fill
to_fill <- c()

for(i in 1:5) {
  
  # Goal: put each x in the to_fill vector
  x <- i + 1
  to_fill <- append(to_fill, x)
  
}
```

```{r}
# the vector i want to fill 
to_fill2 <- rep(NA, 5)

for(i in 1:5) {
  
  # filling it with i + 1
  to_fill2[i] <- i + 1
  
}
```

```{r}
to_fill3 <- seq(from = 2, to = 6, by = 1)
```

# Functions

* allow you to use the same code chunk repeatedly 
   - if you want to do the same thing to a lot of objects, functions are good news!
* lets you generalize your workflow
* makes your code more readable
* you'll need it for assignments in this class!

```{r, eval = FALSE}
function_name <- function(param1, param2, param3, ...) {
  
  # what you want your function to do 
  # suppose your function calculates x based on some parameters
  # x <- _____
  
  
  return(x)
  
}
```


Let's say we want a function that takes in a number, squares it, and then adds 2 to it.

```{r}
square_plus_2 <- function(val) {
  
  y <- (val)^2 + 2
  return(y)
  
}

square_plus_2(5)
square_plus_2(c(1,2,3,4))
```

```{r, eval = FALSE}
# this doesn't work! think about why
square_plus_2("hello")
```

### Improved version:

```{r}
square_plus_2_new <- function(number) {
  
  # check if it's numeric
  if(is.numeric(number) == FALSE) {
    
    print("ERROR!!!!!")
    
  } else {
    
    # calculate and return desired value
    y <- number^2 + 2
    return(y)
    
  }
  
}

square_plus_2_new("hello")
square_plus_2_new(5)
square_plus_2_new(c(1,2,3,4))
```

### Function with iteration

This is just a function with a for (or while) loop inside.

```{r}
my_sum <- function(x, starting_sum = 0) {
  
  # initialize total to be 0 (aka starting_sum) 
  total <- starting_sum
  
  # iterate through the length of x
  for(i in 1:length(x)) {
    
    # increase total by adding x_i onto it
    total <- total + x[i]
    
  }
  
  # return total
  return(total)
  
}

my_sum(x = 1:5)
my_sum(x = 1:3)
```


